---
title: "logiSense Test"
format: html
---

```{r}
library(devtools)
devtools::install_github(
"BelinaJang/logiSense",force = TRUE
)
#devtools::install(build_vignettes = TRUE)
library(logiSense)
```

# load data
```{r}
data("test_data")
```

# case 1: continuous by continuous interaction

```{r}
summary(glm(stroke ~ age*avg_glucose_level, data = test_data, family = "binomial"))
```

## logint1
```{r}
logint(formula=stroke ~ age*avg_glucose_level, variable1 = "age", 
           variable2 = "avg_glucose_level", variable1_type = "continuous", variable2_type = "continuous", data = test_data)
```

## logint3
```{r}
logint3(formula=stroke ~ age*avg_glucose_level, variable1 = "age", 
           variable2 = "avg_glucose_level", variable1_type = "continuous", variable2_type = "continuous", data = test_data)
```

# case 2: continuous by categorical interaction

```{r}
summary(glm(stroke ~ work_type*age, data = test_data, family = "binomial"))
```
## example1: logint
```{r}
logint(stroke ~ work_type*age, variable1 = "work_type", variable2 = "age", variable1_type = "categorical", variable2_type = "continuous",data=test_data, sigfig=4)
```
## example1: logint3
```{r}
logint3(stroke ~ work_type*age, variable1 = "work_type", variable2 = "age", variable1_type = "categorical", variable2_type = "continuous",data=test_data, sigfig=4)
```

## example2: logint
```{r}
logint(stroke ~ work_type*age, variable1 = "age", variable2 = "work_type", variable1_type = "continuous", variable2_type = "categorical",data=test_data, sigfig=4)
```

## example2: logint3
```{r}
logint3(stroke ~ work_type*age, variable1 = "age", variable2 = "work_type", variable1_type = "continuous", variable2_type = "categorical",data=test_data, sigfig=4)
```


# case 3: categorical by categorical interaction

```{r}
summary(glm(stroke ~ work_type*Residence_type, data = test_data, family = "binomial"))
```

## example1: logint
```{r}
logint(stroke ~ work_type*Residence_type, variable1 = "work_type", variable2 = "Residence_type", variable1_type = "categorical", variable2_type = "categorical",data=test_data, sigfig=4)
```

## example1: logint3
```{r}
logint3(stroke ~ work_type*Residence_type, variable1 = "work_type", variable2 = "Residence_type", variable1_type = "categorical", variable2_type = "categorical",data=test_data, sigfig=4)
```

## example2: logint (not in the right order)
```{r}
logint(stroke ~ work_type*Residence_type, variable1 = "Residence_type", variable2 = "work_type", variable1_type = "categorical", variable2_type = "categorical",data=test_data, sigfig=4)
```

## example2: logint3 (not in the right order)
```{r}
logint3(stroke ~ work_type*Residence_type, variable1 = "work_type", variable2 = "Residence_type", variable1_type = "categorical", variable2_type = "categorical",data=test_data, sigfig=4)
```

# error/warning examples
## invalid variable1_type
```{r}
logint3(stroke ~ work_type*Residence_type, variable1 = "work_type", variable2 = "Residence_type", variable1_type = "categoricals", variable2_type = "categorical",data=test_data, sigfig=4)
```
## NA in estimate
```{r}
logint(stroke ~ gender*age, variable1 = "age", variable2 = "gender", variable1_type = "continuous", variable2_type = "categorical",data=test_data, sigfig=4)
```
## interaction term invalid (doesn't exist)
```{r}
formula <- stroke ~ age*work_type + gender
variable1 <- "age"
variable2 <- "gender"
variable1_type <- "continuous"
variable2_type <- "categorical"
data <- test_data
sigfig <- 4


logint(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
logint3(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
```

```{r}
library(tidyverse)

test_data2 <- test_data
test_data2$bmi <- as.numeric(test_data2$bmi)

formula <- stroke ~ age*bmi + avg_glucose_level
variable1 <- "age"
variable2 <- "avg_glucose_level"
variable1_type <- "continuous"
variable2_type <- "continuous"
data <- test_data2
sigfig <- 4


logint(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
logint3(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
```

```{r}
formula <- stroke ~ work_type*gender + avg_glucose_level
variable1 <- "Residence_type"
variable2 <- "gender"
variable1_type <- "categorical"
variable2_type <- "categorical"
data <- test_data
sigfig <- 4


logint(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
logint3(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig)
```
