---
title: "logint_cont"
author: "Jingxuan He"
date: "2024-12-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
test_data <- read.csv(here("data", "test_data.csv"))
```

```{r}
model <- glm(stroke ~ age*avg_glucose_level, family = binomial, data = test_data)
summary(model)
```
# use this version for separate r file case
## should only be run when we know both variables are continuous
```{r}
# continuous x continuous
logint_contbycont <- function(formula, variable1, variable2, data, sigfig=4) {
  continuous_var1 <- variable1
  continuous_var2 <- variable2
  
  cat("continuous_var1:", continuous_var1, "continuous_var2:", continuous_var2, "\n")
  model <- glm(formula, data = data, family = binomial)
  
  variables_list <- as.list(attr(model$terms, "variables"))[-c(1)]
  outcome <- as.character(variables_list[[1]])
  coefficients <- coef(model)
  
  var1_effect <- coefficients[continuous_var1]
  var2_effect <- coefficients[continuous_var2]
  interaction_effect <- coefficients[grep(":", names(coefficients))]
  
  
  # if any coefficients is NA, return warning message
  if (any(is.na(coefficients))) {
    na_variables <- names(coefficients)[is.na(coefficients)]
    warning(paste0(na_variables," has(have) NA estimates. \n Action Required: Consider re-specifying the model or re-examining interaction terms for meaningfulness. \n"))
  }
  
  interaction_effect_name <- names(interaction_effect)
  
  vcov_m <- vcov(model)
  
  sentences <- c(
    paste0(
      "For an observation with ", continuous_var2, "=value, the odds ratio of '", outcome, "' for increasing ", continuous_var1, " by one unit is e^(", signif(var1_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)", ")."
    ),
    paste0("The 95% CI: e^((", signif(var1_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)) \u00b1 (1.96)*sqrt(", signif(vcov_m[continuous_var1, continuous_var1],sigfig), " + ", signif(vcov_m[interaction_effect_name, interaction_effect_name],sigfig), " + 2*(value)*(", signif(vcov_m[continuous_var1, interaction_effect_name],sigfig), "))). \n"
    ),
    paste0("For an observation with ", continuous_var1, "=value, the odds ratio of '", outcome, "' for increasing ", continuous_var2, " by one unit is e^(", signif(var2_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)", ")."
    ),
    paste0("The 95% CI: e^((", signif(var2_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)) \u00b1 (1.96)*sqrt(", signif(vcov_m[continuous_var2, continuous_var2],sigfig), " + ", signif(vcov_m[interaction_effect_name, interaction_effect_name],sigfig), " + 2*(value)*(", signif(vcov_m[continuous_var2, interaction_effect_name],sigfig), "))). \n"
    )
  )
  return(cat(sentences, sep = "\n"))
}
```

logint(formula=stroke ~ age*avg_glucose_level, variable1 = "age", 
           variable2 = "avg_glucose_level", variable1_type = "continuous", variable2_type = "continuous", data = test_data)

above will call below function when both variables are continuous (when using multiple r files)
```{r}
logint_contbycont(formula=stroke ~ age*avg_glucose_level, variable1 = "age", variable2 = "avg_glucose_level", data = test_data, sigfig=4)
```


#input update

```{r}
# continuous x continuous part for logint

logint <- function(formula, variable1, variable2, variable1_type, variable2_type, data, sigfig=4) {
  
  #####################################################
  if (variable1_type == "continuous" & variable2_type == "continuous"){
    #both continuous
    continuous_var1 <- variable1
    continuous_var2 <- variable2
    message("Both variables are continuous. \n")
    message("continuous_var1: ", continuous_var1, ", continuous_var2: ", continuous_var2, "\n")
    
  } else if (variable1_type == "categorical" & variable2_type == "categorical"){
    #both categorical
    categorical_var1 <- variable1
    categorical_var2 <- variable2
    message("Both variables are categorical. \n")
    message("categorical_var1: ", categorical_var1, ", categorical_var2: ", categorical_var2, "\n")
    
  } else if (variable1_type != variable2_type){
    # when two variables are different types, assign each variable to appropriate type variable
    message("Both variables are of different types. \n")
    
    if (variable1_type == "continuous"){
      continuous_var <- variable1
      message("Variable 1 is continuous. \n")
      message("continuous_var: ", continuous_var, "\n")
      } else if (variable1_type == "categorical"){
        categorical_var <- variable1
        message("Variable 1 is categorical. \n")
        message("categorical_var: ", categorical_var, "\n")
      } else {
        stop("variable1_type should be either 'continuous' or 'categorical'.")
      }
    if (variable2_type == "continuous"){
      continuous_var <- variable2
      message("Variable 2 is continuous. \n")
      message("continuous_var: ", continuous_var, "\n")
    } else if (variable2_type == "categorical"){
        categorical_var <- variable2
        message("Variable 2 is categorical. \n")
        message("categorical_var: ", categorical_var, "\n")
    } else {
        stop("variable2_type should be either 'continuous' or 'categorical'.")
    }
 }
  #####################################################
  
  # run below if both variables are continuous
  if (variable1_type=="continuous" & variable2_type=="continuous"){
  
  model <- glm(formula, data = data, family = binomial)
  
  variables_list <- as.list(attr(model$terms, "variables"))[-c(1)]
  outcome <- as.character(variables_list[[1]])
  coefficients <- coef(model)
  
  var1_effect <- coefficients[continuous_var1]
  var2_effect <- coefficients[continuous_var2]
  #interaction_effect <- coefficients[grep(":", names(coefficients))] - wrong
  
    result <- tidy(model)
    
    # error handling added
    result_terms <- result$term

    # find terms that include both continuous and categorical variables from result_terms
    interaction_terms <- coefficients[result_terms[sapply(result_terms, function(term) {all(sapply(c(continuous_var1,continuous_var2), function(var) grepl(var, term)))})]]

    if (length(interaction_terms) == 0){
      stop("No interaction terms between ", continuous_var1, " and ", continuous_var2, " found in the model.")
    }
  
  # if any coefficients is NA, return warning message
  if (any(is.na(coefficients))) {
    na_variables <- names(coefficients)[is.na(coefficients)]
    warning(paste0(na_variables," has(have) NA estimates. \n Action Required: Consider re-specifying the model or re-examining interaction terms for meaningfulness. \n"))
  }
  
  interaction_effect_name <- names(interaction_effect)
  
  vcov_m <- vcov(model)
  
  sentences <- c(
    paste0(
      "For an observation with ", continuous_var2, "=value, the odds ratio of '", outcome, "' for increasing ", continuous_var1, " by one unit is e^(", signif(var1_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)", ")."
    ),
    paste0("The 95% CI: e^((", signif(var1_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)) \u00b1 (1.96)*sqrt(", signif(vcov_m[continuous_var1, continuous_var1],sigfig), " + ", signif(vcov_m[interaction_effect_name, interaction_effect_name],sigfig), " + 2*(value)*(", signif(vcov_m[continuous_var1, interaction_effect_name],sigfig), "))). \n"
    ),
    paste0("For an observation with ", continuous_var1, "=value, the odds ratio of '", outcome, "' for increasing ", continuous_var2, " by one unit is e^(", signif(var2_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)", ")."
    ),
    paste0("The 95% CI: e^((", signif(var2_effect, sigfig), " + (", signif(interaction_effect, sigfig), ")*(value)) \u00b1 (1.96)*sqrt(", signif(vcov_m[continuous_var2, continuous_var2],sigfig), " + ", signif(vcov_m[interaction_effect_name, interaction_effect_name],sigfig), " + 2*(value)*(", signif(vcov_m[continuous_var2, interaction_effect_name],sigfig), "))). \n"
    )
  )
  return(cat(sentences, sep = "\n"))
  }
}
```

# some tests

## example: both variables are continuous
```{r}
logint(formula=stroke ~ age*avg_glucose_level, variable1 = "age", 
           variable2 = "avg_glucose_level", variable1_type = "continuous", variable2_type = "continuous", data = test_data)
```
## example: both variables are categorical
```{r}
logint(formula=stroke ~ age*avg_glucose_level, variable1 = "hypertension", 
           variable2 = "heart_disease", variable1_type = "categorical", variable2_type = "categorical", data = test_data)
```

## example: both variables are of different types
```{r}
logint(formula=stroke ~ age*work_type, variable1 = "work_type", 
           variable2 = "age", variable1_type = "categorical", variable2_type = "continuous", data = test_data)
```

```{r}
logint(formula=stroke ~ age*work_type, variable1 = "age", 
           variable2 = "work_type", variable1_type = "continuous", variable2_type = "categorical", data = test_data)
```

## typo example
```{r}
logint(formula=stroke ~ age*avg_glucose_level, variable1 = "age", 
           variable2 = "work_type", variable1_type = "continuous", variable2_type = "categoricall", data = test_data)
```
