---
title: "functiondraft"
author: "Jingxuan He"
date: "2024-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(broom)
library(here)
```

```{r}
test_data <- read.csv(here("data", "healthcare-dataset-stroke-data.csv"))

glm_model <- glm(stroke ~ gender + age + hypertension + heart_disease +
                   avg_glucose_level + smoking_status, family = binomial, data = test_data)
summary(glm_model)
```


```{r}
# new function: belina
# Revised to handle interaction terms
ilr <- function(formula, variable_interest, variable_type, data) {
  
  model <- glm(formula, data = data, family = binomial)
  
  variables_list <- as.list(attr(model$terms, "variables"))[-c(1)]
  outcome <- as.character(variables_list[[1]])
  cat("Outcome variable: ", outcome, "\n")
  
  result <- tidy(model, conf.int = TRUE, exponentiate = TRUE)
  result_terms <- result$term
  
  interpretation <- ""

  # check if formula contains *
  if (grepl("*", formula)) {
    # Split the interaction term into component variables
    var_list <- unlist(strsplit(variable_interest, "\\*"))
    
    # Split the variable types
    var_type_list <- unlist(strsplit(variable_type, "\\*"))
    
    # Find terms that include all variables in var_list
    result_term_interest <- result_terms[sapply(result_terms, function(term) {
      all(sapply(var_list, function(var) grepl(var, term)))
    })]
    
    # filter result for interaction terms
    interaction_terms <- result[result$term %in% result_term_interest, ]
    
    if (nrow(interaction_terms) > 0) {
      interpretation <- paste0(interpretation, "\nInteraction Effects:\n")
      
      for (i in 1:nrow(interaction_terms)) {
        interaction_term <- interaction_terms$term[i]
        estimate <- interaction_terms$estimate[i]
        ci_lower <- interaction_terms$conf.low[i]
        ci_upper <- interaction_terms$conf.high[i]
        p_value <- interaction_terms$p.value[i]
        
        ci_lower <- ifelse(is.na(ci_lower) | is.infinite(ci_lower), "NA", signif(ci_lower, 4))
        ci_upper <- ifelse(is.na(ci_upper) | is.infinite(ci_upper), "NA", signif(ci_upper, 4))
        
        # skip if estimate is NA
        if (is.na(estimate)) {
          next
        } else {
          p_interpretation <- if (p_value < 0.05) {
            paste0("This interaction is statistically significant at 5% significance level (p-value: ", signif(p_value, 4), ").")
          } else {
            paste0("This interaction is not statistically significant at 5% significance level (p-value: ", signif(p_value, 4), ").")
        }
          
        #interpretation <- paste0(
        #  interpretation,
        #  "The interaction term '", interaction_term, 
        #  "' modifies the effect of '",var_list , "' on '", outcome, 
        #  ".' The odds are multiplied by ", signif(estimate, 4), 
        #  " (95% CI: ", ci_lower, " - ", ci_upper, "). ",
        #  p_interpretation, "\n"
        #)
        
        # loop it thru var_list
        for (i in 1:length(var_list)) {
          if (var_type_list[i] == "categorical") {
            data[[var_list[i]]] <- factor(data[[var_list[i]]])
            levels <- levels(data[[var_list[i]]])
            reference_level <- levels[1]
            # drop all var names in var_list and ":" from interaction_term
            comparison_level <- gsub(paste(var_list, collapse = "|"), "", gsub(":", "", interaction_term))
            
            interpretation <- paste0(
              interpretation,"\n",
              "The interaction term: ", interaction_term, "\n",
              "var_list[i]: ", var_list[i], "\n",
              "  reference_level: ", reference_level, "\n",
              "  comparison_level: ", comparison_level, "\n",
              "  var_type_list[i]", var_type_list[i]
            )
            
          } else if (var_type_list[i] == "continuous") {
            interpretation <- paste0(
              interpretation, "\n",
              "The interaction term: ", interaction_term, "\n",
              "var_list[i]: ", var_list[i], "\n",
              "  var_type_list[i]", var_type_list[i]
            )
          } else {
            stop("Variable type must be either 'continuous' or 'categorical.'")
          }
        }
      }
    }
  }
  }
  
  # Continue with the rest of the function for main effects
  else if (variable_type == "continuous") {
    vi_result <- result[result$term == variable_interest, ]
    if (nrow(vi_result) == 0) {
      stop("Variable of interest not found in the model.")
    }
    
    estimate <- vi_result$estimate
    ci_lower <- vi_result$conf.low
    ci_upper <- vi_result$conf.high
    p_value <- vi_result$p.value
    
    p_interpretation <- if (p_value < 0.05) {
      paste0("This result is statistically significant at 5% significance level (p-value: ", signif(p_value, 4), ").")
    } else {
      paste0("This result is not statistically significant at 5% significance level (p-value: ", signif(p_value, 4), ").")
    }
    
    interpretation <- paste0(
      interpretation,
      "For each one-unit increase in '", variable_interest, ",' the odds of '", outcome,
      "' are multiplied by ", signif(estimate, 4),
      " (95% CI: ", signif(ci_lower, 4), " - ", signif(ci_upper, 4), "). ",
      p_interpretation, "\n"
    )
  } else if (variable_type == "categorical") {
    data[[variable_interest]] <- factor(data[[variable_interest]])
    levels <- levels(data[[variable_interest]])
    reference_level <- levels[1]
    cat("Reference level: ", reference_level, "\n")
    
    for (comparison_level in levels[-1]) {
      term_name <- paste0(variable_interest, comparison_level)
      vi_result <- result[result$term == term_name, ]
      if (nrow(vi_result) > 0) {
        estimate <- vi_result$estimate
        ci_lower <- vi_result$conf.low
        ci_upper <- vi_result$conf.high
        p_value <- vi_result$p.value
        
        ci_lower <- ifelse(is.na(ci_lower) | is.infinite(ci_lower), "NA", signif(ci_lower, 4))
        ci_upper <- ifelse(is.na(ci_upper) | is.infinite(ci_upper), "NA", signif(ci_upper, 4))
        
        p_interpretation <- if (p_value < 0.05) {
          paste0("This result is statistically significant at 5% significance level (p-value: ", signif(p_value, 4), "). ")
        } else {
          paste0("This result is not statistically significant at 5% significance level (p-value: ", signif(p_value, 4), "). ")
        }
        
        interpretation <- paste0(
          interpretation,
          "Compared to the reference level '", reference_level, "' of '",
          variable_interest, 
          ",' the odds of '", outcome, "' for the level '", comparison_level, 
          "' are multiplied by ", signif(estimate, 4),
          " (95% CI: ", ci_lower, " - ", ci_upper, "). ",
          p_interpretation, "\n"
        )
      }
    }
  } else {
    stop("Variable type must be either 'continuous' or 'categorical.'")
  }
  
  return(cat(interpretation))
}

```

```{r warning=FALSE}
# categorical test
result <- ilr(formula=stroke~gender+age+hypertension+heart_disease+avg_glucose_level+smoking_status,data=test_data, variable_interest="gender", variable_type="categorical")

result
```

```{r warning=FALSE}
# continuous test
result <- ilr(stroke~gender+age+hypertension+heart_disease+avg_glucose_level+smoking_status,data=test_data, variable_interest="age", variable_type="continuous")

result
```

```{r}
# double checking!
glm(stroke~gender+age+hypertension+heart_disease+avg_glucose_level+smoking_status,data=test_data, family=binomial)
```

```{r warning=FALSE}
#interaction test
#ilr(formula=stroke~gender*age,data=test_data, variable_interest="gender", variable_type="categorical")
#ilr(formula=stroke~gender*age,data=test_data, variable_interest="age", variable_type="continuous")
ilr(formula=stroke~gender*age,data=test_data, variable_interest="gender*age", variable_type="categorical*continuous")

formula=stroke~gender*age
data=test_data
variable_interest="gender*age"
```
NEED updates to extract reference level like this <- check categorical part for reference
Interaction Effects:
The interaction term 'genderMale:age' modifies the effect of 'gender' for the level "Male" on 'stroke.' The odds are multiplied by 1.012 (95% CI: 0.9919 - 1.033). This interaction is not statistically significant at 5% significance level (p-value: 0.2539).
 
Interaction Effects:
The interaction term 'genderMale:age' modifies the effect of 'age' on 'stroke.' The odds are multiplied by 1.012 (95% CI: 0.9919 - 1.033). This interaction is not statistically significant at 5% significance level (p-value: 0.2539).
```{r}
# verify
glmfit <- summary(glm(stroke~gender*age,data=test_data,family=binomial))
glmfit

int_male <- coef(glmfit)[5][1]
int_other <- coef(glmfit)[6][1]

cat("interaction OR male: ", exp(int_male), "\ninteraction OR other: ", exp(int_other))
```

ref
genderMale:age -> 1.012 (exp(0.011678))
genderMale -> 0.515 exp(-0.664)

For males, the effect of age on the log-odds of stroke is 0.0702 (main effect of age) + 0.0117 (interaction term) = 0.0819.

For females (the reference group), the effect of age on the log-odds of stroke is 0.0702.

This means the impact of age on stroke is slightly stronger for males than for females, but the interaction term is not statistically significant (p = 0.254), so this difference may not be meaningful.
